---
title: "Getting started with conCultuR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with conCultuR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`conCultuR` turns digital traces into auditable biodiversity-attention metrics. The basic workflow is dictionary construction, query ambiguity audit, data collection or simulation, platform standardisation, attention-index estimation, and attention-gap interpretation.

The offline workflow is designed to run without internet access.

```{r}
library(conCultuR)
ex <- con_culturomics_example_data()
dict <- score_ambiguity(ex$dictionary)
head(audit_queries(dict))
```

```{r}
traces <- collect_attention(
  dict,
  platforms = c("google", "wikipedia", "reddit"),
  from = "2024-01-01",
  to = "2024-02-15",
  mock = TRUE
)
traces_std <- standardise_attention(traces)
idx <- attention_index(traces_std, min_platforms = 1)
head(idx)
```

```{r}
gap <- attention_gap(idx, ex$conservation_status)
gap
```

The attention gap is a communication metric. It is not a substitute for ecological risk assessment, population monitoring, or conservation priority setting.
